---
title: "Untitled1"
format: html
---

```{r}
#| message: false

library(dplyr)
library(tidyr)
library(readr)
```

# Data Sources

## Read in our data

```{r}
#| message: false

catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

```{r}
## Prints the column names of my data frame
colnames(catch_original)

## First 6 lines of the data frame
head(catch_original)

## Summary of each column of data
summary(catch_original)

## Prints unique values in a column (in this case, the region)
unique(catch_original$Region)

## Opens data frame in its own tab to see each row and column of the data (do in console)
#View(catch_original)
```

## Explore

# Cleaning & Reshaping

## Using the select () function
```{r}
# Select() to keep the columns that want
catch_data <- catch_original %>%
    select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
head(catch_data)

# to drop columns that you dont want (-)
catch_data <- catch_original %>%
    select(-All,-notesRegCode)
```

## check colum types
```{r}
glimpse(catch_data)
```
## Changing column contents with mutate()
```{r}
# change Chinook column to numeric
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))
```

```{r}
is.na(catch_clean$Chinook)

# get the location of the NA value
i <- which(is.na(catch_clean$Chinook))
i

# look at the row in the data 
catch_data[i,]

# use mutate to fix
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(condition = Chinook == "I",
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))

head(catch_clean)


```
## Changing file shape using pivot_longer() and pivot_wider()

```{r}
catch_long <- catch_clean %>% 
    #pivot longer all columns except Region and Year
    pivot_longer(
        cols = -c(Region, Year),
        names_to = "species",
        values_to = "catch"
    )

head(catch_long)

catch_wide <- catch_long %>%
    pivot_wider(names_from = species,
                values_from = catch)

head(catch_wide)
```
## rename column with rename()
```{r}
catch_long <- catch_long %>%
    rename(catch_thousands = catch)

head(catch_long)
```
## adding columns using mutate
```{r}
catch_long <- catch_long %>%
    mutate(catch = catch_thousands * 1000)

head(catch_long)

catch_long <- catch_long %>%
    mutate(catch = catch_thousands * 1000) %>%
    select(-catch_thousands)

head(catch_long)
```
## summary statistics using group_by() and summarize()
```{r}
unique(catch_original$Region)
```
### split, apply, and comibine
```{r}
mean_region <- catch_long %>%
    group_by(Region) %>%
    summarize(mean_catch = mean(catch))

head(mean_region)


```
### split-Apply-Combine strategy
```{r}
n_region <- catch_long %>%
    group_by(Region) %>%
    summarize(n = n())

head(n_region)
```

```{r}
## for example:
catch_year_sp <- catch_long %>%
    group_by(Year, species) %>%
    summarize(total_year = sum(catch, na.rm = T))
```


## Filtering rows using filter() 
```{r}
sse_catch <- catch_long %>%
    filter(Region == "SSE")

head(sse_catch)
```
# try filtering where catch column is over 1 million fish
```{r}
catch_million <- catch_long %>%
  filter(catch> 1000000)
head(catch_million)
chinook_sse <- catch_long %>%
  filter(Region == "SSE" & species == "Chinook")
head(chinook_sse)

```
## Sorting data with arrange()
```{r}
mean_region <- catch_long %>%
    group_by(Region) %>%
    summarize(mean_catch = mean(catch)) %>%
    arrange(mean_catch)

head(mean_region)
```
### use desc() order 
```{r}
mean_region <- catch_long %>%
    group_by(Region) %>%
    summarize(mean_catch = mean(catch)) %>%
    arrange(desc(mean_catch))

head(mean_region)
```

## Combining datasets together with join()
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```
### column names do not exactly match
```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```
### rename
```{r}
region_defs <- region_defs %>% 
  rename(Region = code, Region_Name = mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = c("Region"))

head(catch_joined)
```


## split and combine values in columns 

### Splitting a column using separate() and unite()
```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAU-039"))

split_df <- sites_df %>%
    separate(site, c("island", "site_number"), sep = "-")
head(split_df)

```
```{r}
dates_df <- data.frame(
    year = c("1930",
             "1930",
             "1930"),
    month = c("12",
              "12",
              "12"),
    day = c("14",
            "15",
            "16")
)

dates_df %>%
    unite(date, year, month, day, sep = "-")
head(dates_df)

dates_unite <- dates_df %>%
  unite(col = date, year, month, day, sep = "-")
head(dates_unite)

```
```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), 
               names_to = "species", 
               values_to = "catch") %>%
  mutate(catch = catch * 1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))

head(mean_region)
```
## Write out the results with readr::write_csv()
```{r}
library(here) 
write_csv(mean_region, here::here("data/mean_catch_by_region.csv"))
```








