---
title: "Introduction to Quarto"
format: html
bibliography: references.bib
---

Markdown formatting

# This is a large title

## This is a smaller title

##### this is an even smaller title

*text* **bold**

[@oke2020]hsdlkjfahljsdfjkladfjl

#1. Introduction

##a. About the data

-   water sample data from Arctic Ocean
-   Data downloaded from the Arctic Data Center: [Tweedie ADC data](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

##b. Setup

#library(readr)

#library(here)

##c. Read in data

```{r}
bg_chem <- read_csv(here::here("data/BGchem2008data.csv"))

```

#2. Analysis

##a. Calculate summary statistics

```{r, echo=FALSE}
#| eval: false
#| echo: false
## Prints the column names of my data frame
colnames(bg_chem)

## General structure of the data frame - shows class of each column
str(bg_chem)

## First 6 lines of the data frame
head(bg_chem)

## Summary of each column of data
summary(bg_chem)

## Prints unique values in a column (in this case Date)
unique(bg_chem$Date)
```

-   

##b. Calculate mean Redfield ratio

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
                
```

##c. Plot Redfield ratio

```{r}
ratio <- (nitrate + nitrite + amm) / phos

```

#3. Conclusion
